Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-03-27 RMS/corrections.tex
up to modeling the hardware

@@

@@

@@

@@

@@

Another concern is that the significance of the paper could be much
better if the paper addressed the problem of modelling RMS in a more
general way. Instead, the paper takes the form of a case study, where
the formal modelling of one specific RMS implementation is presented,
rather than a general method for modelling and analyzing RMS
implementations.  Also, the discussion about soundness and
completeness of the analysis (section V.C) is not fully developed and,
once again, it refers to the specific model rather than being a
general result.

@@

Detailed comments

Section II.B \\ In the definition of IR, it is not clear what s and s'
are. I guess they are states, but the concept of state was not
introduced. \\ In the definition of a class, the authors should
specify that a class includes a collection of rules, otherwise this
fact may be missed.

Section III \\ Page 2 right column, line -24 The implementation is
shown as schedule() in Figure 1 -> The pseudocode of schedule() is
shown in Figure 1

Section IV.A \\ I don't understand the meaning of [ctor] in the
definition of some operations (maybe because I have no specific
background on Maude). I could not find the explanation of this writing
in the introduction to Maude. \\ For the stack sort, the authors
mention operations push, pop and peek, but then you don't write the
definitions of these operations. Instead, they write the definitions
of bottom and \#, which are not explained in the text. This part should
be made clearer. \\ Also, the meaning of NzNat is not explained.

Section V.C \\ In the statement of theorem 1, the definition of
time-robust real-time rewrite theory is missing. Also, the meaning of
tick-stabilizing atomic propositions is not defined. \\ This makes the
section not self-contained. \\ The proof of theorem 2 is missing. The
authors just indicate how the proof could be developed, but they do
not develop it. A possibility would be to point to a document where
this proof has been developed.

@@

1. Rate-monotonic scheduling is a very simple scheduling and has been
sufficiently invesigated in the field of embedded system, the authors
claim the innovation of this paper as the implementation using
real-time maude, however, I cannot see any special point (or
advantage) when using this real-time maude in modelling RMS.

2. RMS has many different realizatio methods, no comparison is given
to justify the strong point of real-time maude when applied to RMS.

@@

@@

@@

@@

>>>>

ANSWER: TODO

@@

ANSWER: The detailed proof is now included in the appendix. It is true
that readers may be interested in the detailed proof of Theorem~2 to
feel convinced of the completeness of our approach. However, the proof
needs a bit more theoretical background about rewriting logic and more
details of our model, which we have simplified in the submission to
ease the understanding of our approach. We would prefer to refer the
detailed proof to a non-anonymous technical report, if possible, in
the final version.

@@

ANSWER: TODO. To revise the modeling section.

@@

ANSWER: The related work section is now improved. More comparisons
with the theoretical approaches and with [24,25] are added. The
differences with [24,25] in the models and in the way of modeling are
mainly discussed. On the other hand, it is a bit difficult to compare
the efficiency and scalability between [24,25] and our approach. One
reason is that we have a different setting with [24] and [25], because
the objectives are different: we aim at verifying an RMS implemention
from a real-world real-time system, while [24] and [25] targeted the
RMS algorithm itself. Another reason is that [24] and [25] presented
no discussion on the efficiency and scalability. Furthermore, they use
\verb|TMSVL| and an extension of SPIN, respectively, as verification
tools, which are not open-source, making it unable to re-implement
their work.

@@

ANSWER: The statement has been corrected. The work presented in this
submission aims at modeling and verifying a realistic RMS
implementation in an industrial avonic system. It is a piece of real
work in the industry. Even though we believe that our approach could
be applied to other systems as well, it means more to that specific
one. In this sense, ``sufficient'' means that the details described in
our model meet the requirements and expectations from the users and
the engineers.

@@

ANSWER: TODO. to make the paper self-contained.

Another concern is that the significance of the paper could be much
better if the paper addressed the problem of modelling RMS in a more
general way. Instead, the paper takes the form of a case study, where
the formal modelling of one specific RMS implementation is presented,
rather than a general method for modelling and analyzing RMS
implementations.

ANSWER: Yes, we agree that we have two choices to do the work
presented in the paper: one is to develop a general way to model RMS
algorithms or implementations, and then instantiate the general model
to get a concrete one for our target system; the other is to shape an
accurate model of the target system directly. To us, there is a trade
off between these two directions. We chose the latter with some
concerns. By our experience, the engineers, who write the codes, and
those, who are in charge of verifying the system or the codes, are
usually not the same persons. When a verification engineer models a
system, he is actually abstacting the system or the codes. It is
difficult to ensure that the model behaves the same as the system or
the codes.  And it is more difficult to let other engineers and users
believe that the model behaves the same as the system or the
codes. Furthermore, the developers and users hope that the model
behaves not only the same as what they had in mind, but also the same
as what is written in the codes. A way out is to make the model not
only behave like the codes, but also ``look'' like the codes. This
requires the model to be as accurate as possible. It makes the model
specific, but improves the engineers' confidence in the model and the
verification results. That is also why we emphasize the
correspondences between the functions in the model and the lines in
the pseudocode, when we introduce our model in Section~IV. On the
other hand, we are also interested in the first choice. This will be
the future work and we would try to get a better balance between them.

Also, the discussion about soundness and
completeness of the analysis (section V.C) is not fully developed and,
once again, it refers to the specific model rather than being a
general result.

ANSWER: The section about soundness and completeness is now made more
self-contained. The detailed proof of Theorem~2 is included in the
appendix. On the other hand, since we chose the accurate way to
achieve the model as discussed above, yes, the proof is a specific
one.

@@

Detailed comments
\begin{itemize}
\item
Section II.B

- In the definition of $IR$, it is not clear what $s$
and $s'$ are. I guess they are states, but the concept of state was
not introduced.

ANSWER: TODO

- In the definition of a class, the authors should
specify that a class includes a collection of rules, otherwise this
fact may be missed.

ANSWER: TODO

\item
Section III

- Page 2 right column, line -24: The implementation is shown as
schedule() in Figure 1 $\rightarrow$ The pseudocode of schedule() is
shown in Figure 1

ANSWER: It is done.

\item
Section IV.A

- I don't understand the meaning of [ctor] in the definition of some
operations (maybe because I have no specific background on Maude). I
could not find the explanation of this writing in the introduction to
Maude.

ANSWER: TODO

- For the stack sort, the authors mention operations push, pop and
peek, but then you don't write the definitions of these
operations. Instead, they write the definitions of bottom and \#,
which are not explained in the text. This part should be made clearer.

ANSWER: TODO

- Also, the meaning of NzNat is not explained.

ANSWER: TODO

\item
Section V.C

- In the statement of theorem 1, the definition of time-robust
real-time rewrite theory is missing. Also, the meaning of
tick-stabilizing atomic propositions is not defined. This makes the
section not self-contained.

ANSWER: TODO

- The proof of theorem 2 is missing. The authors just indicate how the
proof could be developed, but they do not develop it. A possibility
would be to point to a document where this proof has been developed.

ANSWER: The detailed proof is now included in the appendix.  The
detailed proof requires a bit more theoretical background about
rewriting logic and more details of our model, which we tried to
simplify in the submission to ease the understanding of our
approach. We would prefer to refer the detailed proof to a
non-anonymous technical report, if possible, in the final version.
\end{itemize}

@@

1. Rate-monotonic scheduling is a very simple scheduling and has been
sufficiently investigated in the field of embedded system, the authors
claim the innovation of this paper as the implementation using
real-time maude, however, I cannot see any special point (or
advantage) when using this real-time maude in modelling RMS.

ANSWER: We agree that RMS is a very simple scheduling and is the
optimum at the same time among the fixed-prority ones, making it
widely applied and sufficiently investigated. Besides the ideal
setting, the schedulability of RMS was also investigated under
different kinds of implementations with consideration about overhead,
as we discuss in the related work section (Section VI). Despite the
reason that the existing results do not fit our target system using
RMS, the most important motivation to develop the work presented is
that, we aim at verifying the system not only for the schedulability,
but also for the correctness. That is, we want to verify that the
target system does schedule the tasks under the RMS algorithm. The
work presented should be seen as a verification problem, instead of a
schedulability test.

2. RMS has many different realization methods, no comparison is given
to justify the strong point of real-time maude when applied to RMS.

ANSWER: The comparison is given in the related work section (Section
VI). We would say that we are not implementing RMS algorithm in
Real-Time Maude. In fact, we are using Real-Time Maude to model an
realistic implementation of RMS from an industrial avionic
system. There is no theoretical reason telling us that we have to use
Real-Time Maude to model the RMS implementation. However, the existing
verification work as far as we know cannot fit our needs, as discussed
in Section VI. Therefore, we developed the work in the paper, as a
piece of real work in the industry.

@@

ANSWER: As an industrial application, we are not trying to make the
problem as difficult as possible. Instead, we are trying to simplify
the model so that the existing tools are able to solve, while the
target system satisfies the assumptions on the model so that the
verification results are trustworthy. On the other hand, the
assumptions are not so rigid. For example, the mask mechanism
considered as hardware details makes the job initiation times of task
$\tau_i$ possibly not equal to $kT_i$, implying that the deadlines of
the jobs may be not equal to $(k+1)T_i$ (see assumption A1'). Another
example is that non-determinism exists in our model, as discussed in
Section VI.

@@

ANSWER: Yes, we agree that this is an application of rewriting
logic. In this paper, we do apply the existing theories and techniques
of rewriting logic to solve a real industrial problem, which has not
yet been investigated by this sort of techniques, achieving an
industrial contribution.

@@

ANSWER: We do not agree that we should illustrate the mechanism behind
the model checking commands, since the model checker is not part of
our contributions in the paper.

@@

ANSWER: ?

### 2016-03-27 RMS/submission.tex
up to modeling the hardware

In this paper, we use \emph{Real-Time Maude}, a \emph{rewriting}-based
formal modeling language and analysis tool for real-time systems, to
model a realistic implementation of RMS, which serves as a simplified
operating system within an avionic control system, and then verify
several desired properties. Based on a realistic implementation, our
model extends the standard RMS model proposed
in~\cite{DBLP:journals/jacm/LiuL73}, by considering overhead and other
details of the hardware platform.

@@

\subsubsection{Specification}
Real-Time Maude models systems as \emph{modules}. A module specifies a
\emph{real-time rewrite theory} ${\cal R} = (\Sigma, E\cup A , IR,
TR)$, where:
\begin{itemize}
\item $\Sigma$ is an algebraic \emph{signature}, that is, a set of
  declarations of \emph{sorts}, \emph{subsorts} and \emph{function
    symbols}. The function symbols are allowed to be mixfix.
\item $(\Sigma, E\cup A)$ is a \emph{membership equational logic
  theory}, with $E$ a set of conditional \emph{equations} and
  \emph{memberships} on $\Sigma$, and $A$ a set of equational axioms
  such as associativity, commutativity and identity.  $(\Sigma, E\cup
  A)$ models the system's ``static'' states as an algebraic data type,
  and is equipped with a built-in specification of a sort \verb|Time|.
\item $IR$ is a set of \emph{labeled conditional rewrite rules}
  specifying the system's local transitions. Each rule has the form
  $[l]~:~s\rightarrow s'\mbox{ \textbf{if} }\bigwedge^n_{j=1}cond_j$,
  where each $cond_j$ is an equality $u_j=v_j$ and $l$ is a
  \emph{label}. Such a rule specifies an \emph{instantaneous
    transition}, without consuming time, from an instance of $s$ to
  the corresponding instance of $s'$, \emph{provided} the conditions
  hold.
\item $TR$ is a set of (\emph{labeled}) \emph{tick rules} of the form
  $[l]~:~\{s\}\rightarrow\{s'\} \mbox{ \textbf{in time} }t\mbox{
  \textbf{if} }cond$ that specify \emph{timed transitions}, which
  advances time in the \emph{entire} state $s$ by $t$ time units. $IR$
  and $TR$ together model the ``dynamic'' behaviors of the system.
\end{itemize}

In rewriting logic, rewrite rules are applied non-deterministically,
that is, when several rules can be applied on a given $s$, any of them
may be chosen. Hence non-deterministic behaviors can be modeled
naturally in Real-Time Maude.  Real-Time Maude also supports
specifications in \emph{object-oriented} style.  A class declaration
$\texttt{class }C\texttt{ |
}att_1\texttt{:}s_1\texttt{,}\ldots\texttt{,}att_n\texttt{:}s_n$
defines a class $C$ with attributes $att_1$ to $att_n$ of sorts $s_1$
to $s_n$, respectively. An \emph{object} of class $C$ is represented
as a term $\texttt{< } O\texttt{:} C \texttt{ | }
att_1\texttt{:}val_1\texttt{,} \ldots
\texttt{,}att_n\texttt{:}val_n\texttt{ >}$ of sort \verb|Object|,
where $O$, of sort \verb|Oid|, is the object's \emph{identifier}, and
$val_i$ is the value of the attribute $att_i$ with $i\in [1,n]$. A
\emph{subclass} inherits all the attributes and rules of its
superclasses.

@@

\section{The Implementation of RMS}
\label{s:imp}
The implementation is in an industrial avionic control system.
Interrupts would be triggered by, and only by, the clock every $T$,
which we call \emph{interrupt cycle}. When an interrupt request
occurs, if the system is interruptible, i.e. the interrupt mask bit is
cleared, the handler function $schedule()$ will be invoked; otherwise
$schedule()$ will be pending until the interrupt mask bit becomes
cleared.  The implementation is shown as $schedule()$ in
Figure~\ref{a:schedule}, where $taskList$ is the list of periodic
tasks to be scheduled. We assume that the list is in descending order
of priority, and both variables $taskList$ and $timer$ are global. In
this implementation, there is only one kind of interrupt, the period
$T_i$ of each task is a multiple of $T$, and the tasks are
independent, meeting assumption~(A3).

@@

These assumptions make our model different from the standard one. For
instance, with assumption~(A1'), an interrupt request that occurs
during switching will be pending, such that jobs of task $\tau_i$
cannot initiate at $kT_i$. They should wait until switching finishes
and the interrupt mask bit is cleared, which is different
from~(A1). Note that (A3) says that jobs are ready to run at their
initiation times, however, no one can start running at exactly its
initiation time, because scheduling takes time. Under these
assumptions, an example showing the execution of tasks scheduled by
the implementation is depicted in Figure~\ref{sf:rmsimp}. Note that
the second job instance of $\tau_1$ is initiated at time $11$ instead
of $10$, because switching is performed and the interrupt mask bit is
set at time $10$.

@@

In this section, we introduce first how we model the states--the
static aspect--of the system using algebraic data types, then how we
specify the essential behaviors--the dynamic aspect--using rewrite
rules. In particular, modeling of \emph{instantaneous behaviors} would
be explained in Sections~\ref{ss:ir}, \ref{ss:init}
and~\ref{ss:inthandling}, followed by \emph{timed behaviors} in
Section~\ref{ss:timedbehavior}.

\subsection{Basic Data Types}
In our model, tasks are identified by their indexes of sort \verb|Nat|
in the $taskList$. We define a sort \verb|MaybeNat| wrapping
\verb|Nat|s to refer to some task, with \emph{constructor} \verb|some|
followed by a \verb|Nat| $n$ indicating the task indexed $n$, and
\verb|none| for no task:
\begin{verbatim}
  op none : -> MaybeNat [ctor] .
  op some_ : Nat -> MaybeNat [ctor] .
\end{verbatim}

A sort \verb|Stack| is introduced to model the stack of the system,
storing the tasks that are being interrupted, and equipped with
operations \verb|push|, \verb|pop| and \verb|peek| on it.
\begin{verbatim}
  op bottom : -> Stack [ctor] .
  op _#_ : Nat Stack -> Stack [ctor] .
\end{verbatim}

We also need
%a sort \verb|Interval| to denote computation requirements of tasks, and 
a sort \verb|Counter| to record the execution of tasks.  We call
\emph{execution time} the time how long a task has been executing
for. A \verb|Counter| records the execution time and the computation
requirement of a task.
\begin{verbatim}
  op [_/_] : Time Time -> Counter [ctor] .
\end{verbatim}

@@

The global variable $timer$ is reset when it reaches an upper bound
while increasing, which is not shown in detail in
Figure~\ref{a:schedule} but is reasonable. The upper bound is the
least common multiple of the periods of all tasks. A sort \verb|Timer|
is defined:
\begin{verbatim}
  op [_/_] : Nat NzNat -> Timer [ctor] .
\end{verbatim}

@@

A periodic task, which needs to be scheduled, is an object instance of
the subclass \verb|PTask| of \verb|Task| with additional attributes
\verb|priority|, \verb|period| and \verb|status|:
\begin{verbatim}
  class PTask | priority : Nat, period : NzNat, 
                status : Status .
  subclass PTask < Task .
\end{verbatim}
where \verb|Status| is a sort with four constant constructors, same as
in the implementation:
\begin{verbatim}
  ops RUNNING INTERRUPT READY DORMANT 
        : -> Status [ctor] .
\end{verbatim}
The list of periodic tasks, the variable $taskList$ in the
implementation, is modeled as an instance of the sort \verb|TaskList|,
which is a list of \verb|PTask|s\footnote{Following the Maude
  convention, variables would be written in capital letters.}:
\begin{verbatim}
  op null : -> TaskList [ctor] .
  op _::_ : Object TaskList 
              ~> TaskList [ctor] .
  mb (< O:Oid : PTask |> :: L:TaskList) 
       : TaskList .
  mb (error :: L:TaskList) : TaskList .
\end{verbatim}
Periodic tasks are identified by their indexes in the list.

@@

\section{Conclusion}
\label{s:conclusion}
We have formally modeled a realistic implementation of RMS using
Real-Time Maude, a modeling language based on rewriting logic. By
taking into account the overhead of scheduling and switching, and by
modeling some mechanism of the hardware, our model contains sufficient
details to be analyzed for the behaviors of real systems. Two
important properties--schedulability and correctness--are verified by
model checking on our model, within several key scenarios. We
demonstrate the soundness and completeness of our results.

\appendix[Proof of Theorem~\ref{t:main}]
\setcounter{theorem}{0}
\renewcommand{\thetheorem}{\thesection.\arabic{theorem}}
\newcommand{\lrps}[2]{\rightarrow^{#1}_{#2}}

@@

\jx{

  Using a model-based checking approach enhance the ability to check
  other scheduling algorithms.

  About non-determinism: task finishes right at the time interrupt
  comes, theoretical results cannot solve

  The chinese one:

  - the model is ideal

  - it should give a model to each task. And given different task sets, the model should be
  adjusted. But our model need only an instantiation of the task set.

  - the time domain. It uses a global integer to model the time, hence the time dense
  must be discrete. But our model use a generic time domain, which can be instantiated on
  both dense and discrete domain.
}

>>>>

In this paper, we use \emph{Real-Time Maude}, a \emph{rewriting}-based
formal modeling language and analysis tool for real-time systems, to
model a realistic implementation of RMS that serves as a simplified
operating system within an avionic control system, and then verify
several desired properties. Based on a realistic implementation, our
model extends the standard RMS model proposed
in~\cite{DBLP:journals/jacm/LiuL73}, by considering overhead and other
details of the hardware platform.

@@

\subsubsection{Specification}
Real-Time Maude models systems as \emph{modules}. A module specifies a
\emph{real-time rewrite theory} ${\cal R} = (\Sigma, E\cup A , IR,
TR)$, where:
\begin{itemize}
\item $\Sigma$ is an algebraic \emph{signature}, that is, a set of
  declarations of \emph{sorts}, \emph{subsorts} and \emph{function
    symbols}. The function symbols are allowed to be mixfix, in which
  case underscores ``\verb|_|'' indicate the positions of parameters.
  \emph{Terms} are expressions built from function symbols and
  variables.
\item $(\Sigma, E\cup A)$ is a \emph{membership equational logic
  theory}, with $E$ a set of conditional \emph{equations} and
  \emph{memberships} on $\Sigma$, and $A$ a set of equational axioms
  such as associativity, commutativity and identity.  $(\Sigma, E\cup
  A)$ models the system's ``static'' states as terms of some sort, and
  is equipped with a built-in specification of a sort \verb|Time|.
\item $IR$ is a set of \emph{labeled conditional rewrite rules}
  specifying the system's local transitions. Each rule has the form
  $[l]~:~t\rightarrow t'\mbox{ \textbf{if} }\bigwedge^n_{j=1}cond_j$,
  where each $cond_j$ is an equality $u_j=v_j$ and $l$ is a
  \emph{label}, $t,t'$ are terms. Such a rule specifies an
  \emph{instantaneous transition}, without consuming time, from an
  instance of $t$ to the corresponding instance of $t'$,
  \emph{provided} the conditions hold.
\item $TR$ is a set of (\emph{labeled}) \emph{tick rules} of the form
  $[l]~:~\{s\}\rightarrow\{s'\} \mbox{ \textbf{in time} }r\mbox{
  \textbf{if} }cond$ that specify \emph{timed transitions}, which
  advances time by $r$ time units in the \emph{entire} state modeled
  by term $s$. $IR$ and $TR$ together model the ``dynamic'' behaviors
  of the system.
\end{itemize}

In rewriting logic, rewrite rules are applied non-deterministically,
that is, when several rules can be applied on a given term $t$, any of
them may be chosen. Hence non-deterministic behaviors can be modeled
naturally in Real-Time Maude.  Real-Time Maude also supports
specifications in \emph{object-oriented} style.  A class declaration
$\texttt{class }C\texttt{ |
}att_1\texttt{:}s_1\texttt{,}\ldots\texttt{,}att_n\texttt{:}s_n$
defines a class $C$ with attributes $att_1$ to $att_n$ of sorts $s_1$
to $s_n$, respectively. An \emph{object} of class $C$ is represented
as a term $\texttt{< } O\texttt{:} C \texttt{ | }
att_1\texttt{:}val_1\texttt{,} \ldots
\texttt{,}att_n\texttt{:}val_n\texttt{ >}$ of sort \verb|Object|,
where $O$, of sort \verb|Oid|, is the object's \emph{identifier}, and
$val_i$ is the value of the attribute $att_i$ with $i\in [1,n]$. Rules
can be defined on a given class. A \emph{subclass} inherits all the
attributes and rules of its superclasses.

@@

\section{The Implementation of RMS}
\label{s:imp}
The implementation written in C is from an industrial avionic control
system.  Interrupts would be triggered by, and only by, the clock
every $T$, which we call \emph{interrupt cycle}. When an interrupt
request occurs, if the system is interruptible, i.e. the interrupt
mask bit is cleared, the handler function $schedule()$ will be
invoked; otherwise $schedule()$ will be pending until the interrupt
mask bit becomes cleared.  The pseudocode of $schedule()$ is shown in
Figure~\ref{a:schedule}, where $taskList$ is the list of periodic
tasks to be scheduled. We assume that the list is in descending order
of priority, and both variables $taskList$ and $timer$ are global. In
this implementation, there is only one kind of interrupt, the period
$T_i$ of each task is a multiple of $T$, and the tasks are
independent, meeting assumption~(A3).

@@

These assumptions make our model different from the standard one. For
instance, with assumption~(A1'), an interrupt request that occurs
during switching will be pending, such that jobs of task $\tau_i$
cannot initiate at $kT_i$. They should wait until switching finishes
and the interrupt mask bit is cleared, which is different
from~(A1). Note that (A3) says that jobs are ready to run at their
initiation times, however, no one can start running at exactly its
initiation time, because scheduling takes time. Under these
assumptions, an example showing the execution of tasks scheduled by
the implementation is depicted in Figure~\ref{sf:rmsimp}. Note that
the second job instance of $\tau_1$ is initiated at time $11$ instead
of $10$, because switching is performed and the interrupt mask bit is
set at time $10$, exemplifying differences between (A1') and (A1).

@@

In this section, we introduce first how we model the states--the
static aspect--of the system using terms of given sorts, or called
data types, then how we specify the essential behaviors--the dynamic
aspect--using rewrite rules. In particular, modeling of
\emph{instantaneous behaviors} would be explained in
Sections~\ref{ss:ir}, \ref{ss:init} and~\ref{ss:inthandling}, followed
by \emph{timed behaviors} in Section~\ref{ss:timedbehavior}.

\subsection{Basic Data Types}
In our model, tasks are identified by their indexes of sort \verb|Nat|
in the $taskList$. We define a sort \verb|MaybeNat| wrapping
\verb|Nat|s to refer to some task, with \emph{constructor} \verb|some|
followed by a \verb|Nat| $n$ indicating the task indexed $n$, and
\verb|none| for no task:
\begin{verbatim}
  op none : -> MaybeNat [ctor] .
  op some_ : Nat -> MaybeNat [ctor] .
\end{verbatim}
where the keyword \verb|ctor| denotes the corresponding function
symbol to be a constructor.

A sort \verb|Stack| is introduced to model the stack of the system,
storing the tasks that are being interrupted, and equipped with
operations \verb|push|, \verb|pop| and \verb|peek| on it.
\hide{
\begin{verbatim}
  op bottom : -> Stack [ctor] .
  op _#_ : Nat Stack -> Stack [ctor] .
\end{verbatim}
}

We also need
%a sort \verb|Interval| to denote computation requirements of tasks, and 
a sort \verb|Counter| to record the execution of tasks.  We call
\emph{execution time} the time how long a task has been executing
for. A \verb|Counter| records the execution time and the computation
requirement of a task.
\hide{
\begin{verbatim}
  op [_/_] : Time Time -> Counter [ctor] .
\end{verbatim}
}

@@

The global variable $timer$ is reset when it reaches an upper bound
while increasing, which is not shown in detail in
Figure~\ref{a:schedule} but is reasonable. The upper bound is the
least common multiple of the periods of all tasks. A sort \verb|Timer|
is defined to model the $timer$.
\hide{
\begin{verbatim}
  op [_/_] : Nat NzNat -> Timer [ctor] .
\end{verbatim}
}

@@

A periodic task, which needs to be scheduled, is an object instance of
the subclass \verb|PTask| of \verb|Task| with additional attributes
\verb|priority|, \verb|period| and \verb|status|:
%  class PTask | priority : Nat, period : NzNat, 
\begin{verbatim}
  class PTask | priority : Nat, period : Nat, 
                status : Status .
  subclass PTask < Task .
\end{verbatim}
where \verb|Status| is a sort with four constant constructors
\verb|RUNNING|, \verb|INTERRUPT|, \verb|READY| and \verb|DORMANT|,
same as in the implementation.
\hide{
\begin{verbatim}
  ops RUNNING INTERRUPT READY DORMANT 
        : -> Status [ctor] .
\end{verbatim}
}

The list of periodic tasks, the variable $taskList$ in the
implementation, is modeled as an instance of sort \verb|TaskList|,
which is a list of \verb|PTask|s and/or \verb|error|s.
\hide{
\begin{verbatim}
  op null : -> TaskList [ctor] .
  op _::_ : Object TaskList 
              ~> TaskList [ctor] .
  mb (< O:Oid : PTask |> :: L:TaskList) 
       : TaskList .
  mb (error :: L:TaskList) : TaskList .
\end{verbatim}
}
Periodic tasks are identified by their indexes in the list.

@@

\section{Conclusion}
\label{s:conclusion}
We have formally modeled a realistic implementation of RMS using
Real-Time Maude, a modeling language based on rewriting logic. By
taking into account the overhead of scheduling and switching, and by
modeling some mechanism of the hardware, our model contains sufficient
details to be analyzed for the behaviors of the real target
system. Two important properties--schedulability and correctness--are
verified by model checking on our model, within several key
scenarios. We demonstrate the soundness and completeness of our
results.

\appendix[Proof of Theorem~\ref{t:main}]
\label{a:proof}
\setcounter{theorem}{0}
\renewcommand{\thetheorem}{\thesection.\arabic{theorem}}
\newcommand{\lrps}[2]{\rightarrow^{#1}_{#2}}

@@
